---
title: "name needed!!"
output: 
  rmarkdown::html_vignette:
    toc: TRUE
  github_document:
    toc: TRUE
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 6,
                      fig.height = 8)

library(futile.logger)
library(tidyverse)
library(reshape2)
library(loose.rock)

.Last.value <- flog.layout(layout.format('[~l] ~m'))
.Last.value <- flog.appender(appender.tee('logger.txt'))
theme_set(theme_minimal())

devtools::load_all()
```

```{r generate_md, eval=FALSE, include=FALSE}
rmarkdown::render('README.Rmd', output_format = 'all')
```

> Elastic-Net models with additional regularization based on network centrality metrics

## Install

Bioconductor is necessary for the installation of this package.

```{r instal, eval=FALSE}
source("https://bioconductor.org/biocLite.R")
biocLite('averissimo/network.cox', dependencies=TRUE, build_vignettes=FALSE)
library(network.cox)
```

## Citation

...

## Overview

This package extends the `glmnet` r-package with network-based regularization based on features relations. This network can be calculated from the data itself or using external networks to enrich the model.

It adds two new main functions called `network.glmnet` and `network.cv.glmnet` that extend both model inference and model selection via cross-validation with network-based regularization.

There are 3 methods available to use data-dependant methods to generate the netork:

1. Correlation matrix with cutoff;
1. Covariance matrix with cutoff;
<!-- 1. Sparse bayesian networks using `sparsebn` package. -->

Alternatively, the network can be passed as a adjancency matrix or an already calculate metric for each node.

### Example for gaussian models

The example below, shows random datasets being generated and `network.glmnet` new function being called.

```{r gaussian}
# Gaussian
x <- matrix(rnorm(100*20),100,20)
y <- rnorm(100)
fit1 <- network.glmnet(x,y, 'correlation')
```

Inspecting the penalty.factor used from correlation network

```{r}
fit1$penalty.factor
```

Plot the results of the `glmnet` run

```{r}
plot(fit1)
```


The result can be used with all functions available to glmnet objects, such as `predict`, `coef` or plot

```{r predicted}
predicted <- predict(fit1, newx=x[1:10,],s=c(0.01,0.005))
```

```{r pred.table, echo=FALSE}
colnames(predicted) <- c('lambda_0.01', 'lambda_0.005')
flog.info('Observed vs. Predicted', cbind(Observed = y[1:10], predicted), capture = TRUE)
```


It also extends the new methods to the cross validation function with `network.cv.glmnet`

```{r plot_cv}
plot(network.cv.glmnet(x,y, 'covariance'))
```

### Survival Example using RNASeq data

We use an example data from TCGA Adrenocortical Carcinoma project with '92' patients and a reduced RNASeq data. See `MultiAssayExperiment::miniACC` for more information on the data.

There is some pre-processing needed to remove patients with invalid follow-up date or death date:

```{r multi.assay}
# load data
xdata <- MultiAssayExperiment::miniACC

# build valid data with days of last follow up or to event
event.ix <- which(!is.na(xdata$days_to_death))
cens.ix  <- which(!is.na(xdata$days_to_last_followup))
surv_event_time <- array(NA, nrow(xdata@colData))
surv_event_time[event.ix] <- xdata$days_to_death[event.ix]
surv_event_time[cens.ix]  <- xdata$days_to_last_followup[cens.ix]

# Keep only valid individuals
#
# they are valid if they have:
#  - either a follow_up time or event time
#  - a valid vital_status (i.e. not missing)
#  - folloup_time or event_time > 0
valid.ix <- as.vector(!is.na(surv_event_time) & !is.na(xdata$vital_status) & surv_event_time > 0)
ydata <- data.frame(time      = surv_event_time[valid.ix], 
                    status    = xdata$vital_status[valid.ix], 
                    row.names = xdata$patientID[valid.ix])

```

Fitting the survival model using a correlation network with cutoff at 0.6

```{r fit.surv}
# build response object for glmnet
fit3 <- network.cv.glmnet(xdata, ydata, family = 'cox', 
                          network = 'correlation', 
                          experiment.name = 'RNASeq2GeneNorm', 
                          alpha = .7,
                          nlambda = 1000,
                          network.options = network.options.default(cutoff = .6, min.degree = 0.2))
plot(fit3)
```


```{r, eval=FALSE, include=FALSE}
library(STRINGdb)
get_STRING_species(version="10", species_name=NULL) %>% 
  arrange(official_name) %>%
  filter(official_name == 'Homo sapiens')

# downloading Homo sapiens
dir.create('../tmp')
string_db <- STRINGdb$new(version="10", species=9606, score_threshold=0, input_directory=file.path('..', 'tmp'))

tp53 = string_db$mp( "tp53" )
atm = string_db$mp( "atm" )

all.interactions <- string_db$get_interactions(string_db$proteins$protein_external_id)
dim(all.interactions)
colnames(all.interactions)
#
all.interactions[1:15,]
#
ixs <- seq(nrow(all.interactions))
cbind(rowSums(all.interactions[ixs,3:15]), 
      all.interactions[ixs,16], 
      sort(rowSums(all.interactions[ixs,3:15]), index.return = TRUE)$ix, 
      sort(all.interactions[ixs,16], index.return = TRUE)$ix)
```

## Visualization tools

### Survival curves with `draw.kaplan`

This generates Kaplan-Meier survival model based on the coefficients of a Cox model. It separates two groups based on relative risk and display both the curves and results of Log-rank test.

```{r draw.kaplan}
xdata.reduced <- filter.by.experiment(xdata, 'RNASeq2GeneNorm')
ydata.km <- ydata[rownames(xdata.reduced@colData),]
best.model.coef <- coef(fit3, s = 'lambda.min')[,1]
draw.kaplan(best.model.coef, t(assay(xdata[['RNASeq2GeneNorm']])), ydata.km)
```

### Heatmap with results from Hallmarks of cancer

Search the non-zero coefficients in the results and query for known hallmarks of cancer.

```{r hallmarks}
fit3.hallmarks <- hallmarks(names(best.model.coef)[best.model.coef > 0])

melt(fit3.hallmarks$hallmarks, id.vars = 'gene.name') %>%
  filter(value > 0) %>%
  ggplot() + 
    geom_raster(aes(gene.name, variable, fill=value)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

