---
title: "Example for Classification Data -- Breast Invasive Carcinoma"
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document:
    number_sections: yes
    toc: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Example for Classification Data -- Breast Invasive Carcinoma}
  %\VignetteEncoding{UTF-8}
params:
  seed: !r 29221
--- 

## Instalation

```{r, eval=FALSE}
source("https://bioconductor.org/biocLite.R")
biocLite('glmSparseNet')
```

# Required Packages

```{r packages, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(survival)
library(ggfortify)
library(loose.rock)
library(futile.logger)
#
library(glmSparseNet)
#
# Some general options for futile.logger the debugging package
.Last.value <- flog.layout(layout.format('[~l] ~m'))
.Last.value <- loose.rock::show.message(FALSE)
# Setting ggplot2 default theme as minimal
theme_set(ggplot2::theme_minimal())
```

# Load data

The data is loaded from an online curated dataset downloaded from TCGA using `TCGAbiolinks` bioconductor package and processed.

To accelerate the process we use a very reduced dataset down to 107 variables only *(genes)*, which is stored as a data object in this package. However, the procedure to obtain the data manually is described in the following chunk.

```{r data, eval=FALSE}
# download data package with TCGA data
biocLite('https://github.com/averissimo/tcga.data/releases/download/2016.12.15-brca/brca.data_1.0.tar.gz')

# load library
library(brca.data)

# load data
data('fpkm.per.tissue', 'fpkm.per.tissue.barcode', 'clinical', package = 'brca.data')

#
# RNA-Seq data

# tumor tissue
xdata_tumor <- fpkm.per.tissue$primary.solid.tumor
xdata_tumor <- xdata_tumor[,-which(duplicated(getParticipantCode(colnames(fpkm.per.tissue$primary.solid.tumor))))] # removing duplicated samples

# normal tissue
set.seed(params$seed)
xdata_normal <- fpkm.per.tissue$solid.tissue.normal
xdata        <- rbind(t(xdata_tumor),t(xdata_normal))
xdata        <- xdata[, apply(xdata, 2, sd) != 0] # remove genes with sd == 0

# response vector ('1', tumor; '0', normal)
ydata <- c(rep(1,dim(xdata_tumor)[2]),rep(0,dim(xdata_normal)[2]))

# Using only a subset of genes previously selected to keep this example short.

set.seed(params$seed)
small.subset <- c('ENSG00000110448', 'ENSG00000124449', 'ENSG00000103494', 
                  'ENSG00000172053', 'ENSG00000136758', 'ENSG00000051382', 
                  'ENSG00000128655', 'ENSG00000100368', 'ENSG00000115884', 
                  'ENSG00000179363', 'ENSG00000091106', 'ENSG00000178403', 
                  'ENSG00000066422', 'ENSG00000110448', 'ENSG00000124449', 
                  'ENSG00000230445', 'ENSG00000172053', 'ENSG00000185122', 
                   sample(colnames(xdata), 100))
xdata <- xdata[,small.subset]
```

```{r, eval=FALSE, include=FALSE}
# save data in package
sample.BRCA.logisitic <- list(xdata = xdata, ydata = ydata)
devtools::use_data(sample.BRCA.logistic, overwrite = TRUE)
```

```{r, include=FALSE}
xdata <- sample.BRCA.logistic$xdata
ydata <- sample.BRCA.logistic$ydata
```

# Fit models

Fit model model penalizing by the hubs using the cross-validation function by `cv.glmHub`.

```{r fit.show}
fitted <- cv.glmHub(xdata, ydata, 
                    family  = 'binomial',
                    network = 'correlation', 
                    nlambda = 1000)
```

# Results of Cross Validation

Shows the results of `1000` different parameters used to find the optimal value in 10-fold cross-validation. The two vertical dotted lines represent the best model and a model with less variables selected *(genes)*, but within a standard error distance from the best.

```{r results}
plot(fitted)
```

## Coefficients of selected model from Cross-Validation

Taking the best model described by `lambda.min`

```{r show_coefs}
coefs.v <- coef(fitted, s = 'lambda.min')[,1] %>% { .[. != 0]}
coefs.v %>% { 
  data.frame(ensembl.id  = names(.), 
             gene.name   = gene.names(names(.))$external_gene_name, 
             coefficient = .,
             stringsAsFactors = FALSE)
  } %>%
  arrange(gene.name) %>%
  knitr::kable()
```

## Hallmarks of Cancer

```{r hallmarks}
gene.names(names(coefs.v)) %>% { hallmarks(.$external_gene_name)$heatmap }
```

## Accuracy

```{r accuracy, echo=FALSE}
resp <- predict(fitted, s = 'lambda.min', newx = xdata, type = 'response')
flog.info('Misclassified (%d)', sum(abs(ydata - resp) > 0.5))
flog.info('  * False positive: %d', sum(ydata - resp > 0.5))
flog.info('  * False negative: %d', sum(ydata - resp < -0.5))
```

Histogram of predicted response

```{r predict, echo=FALSE}
qplot(resp, bins = 100)
```

Prediction vs. Observed

```{r}
qplot(resp, ydata, color = abs(resp - ydata) < 0.5) + theme(legend.position = 'none')
```

ROC curve

```{r roc, echo=FALSE}
roc_obj <- pROC::roc(ydata, as.vector(resp))

data.frame(TPR = roc_obj$sensitivities, FPR = 1 - roc_obj$specificities) %>%
  ggplot() +geom_line(aes(FPR,TPR), color = 2, size = 1, alpha = 0.7)+
      labs(title= sprintf("ROC curve (AUC = %f)", pROC::auc(roc_obj)), 
           x = "False Positive Rate (1-Specificity)", 
           y = "True Positive Rate (Sensitivity)")
```

