---
title: "Example for Survival Data -- Prostate Adenocarcinoma"
author: "Marta Lopes and André Veríssimo"
date: "`r Sys.Date()`"
output: 
  BiocStyle::html_document:
    number_sections: yes
    toc: true
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Example for Survival Data -- Prostate Adenocarcinoma}
  %\VignetteEncoding{UTF-8}
params:
  seed: !r 2924
--- 

# Instalation

```{r, eval=FALSE}
if (!require("BiocManager")) {
  install.packages("BiocManager")
}
BiocManager::install("glmSparseNet")
```

# Required Packages

```{r packages, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(survival)
library(futile.logger)
library(curatedTCGAData)
library(TCGAutils)
library(MultiAssayExperiment)
#
library(glmSparseNet)
#
# Some general options for futile.logger the debugging package
.Last.value <- flog.layout(layout.format("[~l] ~m"))
.Last.value <- glmSparseNet:::.showMessage(FALSE)
# Setting ggplot2 default theme as minimal
theme_set(ggplot2::theme_minimal())
```

# Load data

The data is loaded from an online curated dataset downloaded from TCGA using 
`curatedTCGAData` bioconductor package and processed.

To accelerate the process we use a very reduced dataset down to around 100 
variables only *(genes)*, which is stored as a data object in this package. 
However, the procedure to obtain the data manually is described in the 
following chunk.

```{r curated_data, include=FALSE}
# chunk not included as it produces to many unnecessary messages
prad <- tryCatch(
  {
    curatedTCGAData(
      diseaseCode = "PRAD",
      assays = "RNASeq2GeneNorm",
      version = "1.1.38",
      dry.run = FALSE
    )
  },
  error = function(err) {
    NULL
  }
)
```

```{r curated_data_non_eval, eval=FALSE}
prad <- curatedTCGAData(
  diseaseCode = "PRAD", assays = "RNASeq2GeneNorm",
  version = "1.1.38", dry.run = FALSE
)
```

Build the survival data from the clinical columns.

* Selects only primary solid tumour samples
* Merge survival times for patients, which have different columns in case they 
are alive or dead.
* Build two matrix objects that fit the data `xdata` and `ydata`

```{r data.show, warning=FALSE, error=FALSE, eval=!is.null(prad)}
# keep only solid tumour (code: 01)
prad.primary.solid.tumor <- TCGAutils::TCGAsplitAssays(prad, "01")
xdata.raw <- t(assay(prad.primary.solid.tumor[[1]]))

# Get survival information
ydata.raw <- colData(prad.primary.solid.tumor) |>
  as.data.frame() |>
  # Find max time between all days (ignoring missings)
  dplyr::rowwise() |>
  dplyr::mutate(
    time = max(days_to_last_followup, days_to_death, na.rm = TRUE)
  ) |>
  # Keep only survival variables and codes
  dplyr::select(patientID, status = vital_status, time) |>
  # Discard individuals with survival time less or equal to 0
  dplyr::filter(!is.na(time) & time > 0) |>
  as.data.frame()

# Set index as the patientID
rownames(ydata.raw) <- ydata.raw$patientID

# keep only features that have standard deviation > 0
xdata.raw <- xdata.raw[
  TCGAbarcode(rownames(xdata.raw)) %in% rownames(ydata.raw),
]
xdata.raw <- xdata.raw[, apply(xdata.raw, 2, sd) != 0] |>
  scale()

# Order ydata the same as assay
ydata.raw <- ydata.raw[TCGAbarcode(rownames(xdata.raw)), ]

set.seed(params$seed)
small.subset <- c(
  geneNames(c(
    "ENSG00000103091", "ENSG00000064787",
    "ENSG00000119915", "ENSG00000120158",
    "ENSG00000114491", "ENSG00000204176",
    "ENSG00000138399"
  ))$external_gene_name,
  sample(colnames(xdata.raw), 100)
) |>
  unique() |>
  sort()

xdata <- xdata.raw[, small.subset[small.subset %in% colnames(xdata.raw)]]
ydata <- ydata.raw |> dplyr::select(time, status)
```

# Fit models

Fit model model penalizing by the hubs using the cross-validation function by 
`cv.glmHub`.

```{r fit, eval=!is.null(prad)}
set.seed(params$seed)
fitted <- cv.glmHub(xdata, Surv(ydata$time, ydata$status),
  family = "cox",
  nlambda = 1000,
  network = "correlation",
  network.options = networkOptions(
    cutoff = .6,
    minDegree = .2
  )
)
```

# Results of Cross Validation

Shows the results of `100` different parameters used to find the optimal value
in 10-fold cross-validation. The two vertical dotted lines represent the best
model and a model with less variables selected *(genes)*, but within a standard 
error distance from the best.

```{r results, eval=!is.null(prad)}
plot(fitted)
```

## Coefficients of selected model from Cross-Validation

Taking the best model described by `lambda.min`

```{r show_coefs, eval=!is.null(prad)}
coefs.v <- Filter(function(.x) .x != 0, coef(fitted, s = "lambda.min")[, 1])
data.frame(
  ensembl.id = names(coefs.v),
  gene.name = geneNames(names(coefs.v))$external_gene_name,
  coefficient = coefs.v,
  stringsAsFactors = FALSE
) |>
  arrange(gene.name) |>
  knitr::kable()
```

## Survival curves and Log rank test

```{r, eval=!is.null(prad)}
separate2GroupsCox(as.vector(coefs.v),
  xdata[, names(coefs.v)],
  ydata,
  plot.title = "Full dataset", legend.outside = FALSE
)
```

# Session Info

```{r sessionInfo}
sessionInfo()
```
